# APPENDICES {.unnumbered}

```{r}
library(patchwork)
plot_raster <- function(x = NULL, lyr = "Average", what = "tmap", log = FALSE){
  library(terra)
  library(tmap)
  library(ggplot2)
  library(tidyterra)
  r <- rast(x=x, lyrs = lyr)
  l.title <- "Density"
  if (log == TRUE){ 
    r <- log(r)
    l.title <- "Density"
  }
  if(what == "tmap"){
  p <- tm_shape(r) + 
    tm_raster(col = lyr, 
              col.scale = tm_scale_continuous(values = "magma"),
              col.legend = tm_legend(title = l.title)) +
    tm_layout(
      legend.outside = TRUE,         # Move legend outside plot
      legend.outside.position = "top", # Place outside at top
      legend.orientation = "landscape")
  }
  if(what == "gg"){
    #rescale raster for plotting
    r <- ifel(r < -8, -8, r)
    mybreaks <- seq(from =max(-8, minmax(r)[1]), to = minmax(r)[2], length.out = 5)
    p <- ggplot() +
      geom_spatraster(data = r) +
      geom_spatraster_contour(data = r, color = "lightgray") +
      # scale_fill_whitebox_c() + 
      # geom_spatraster_contour_filled(data = r) + 
      scale_fill_viridis_c(option = "magma", na.value = "transparent", 
                           name = l.title, 
                           breaks = mybreaks,
                           #n.breaks = 5,
                           labels = function(b) round(exp(b), 3)) + 
      # geom_spatraster_contour(data = r, color = "white") +
      coord_sf(crs = 3338) + 
      #guides(fill=guide_legend(title="Density")) +
      theme(legend.position = "top", 
            legend.key.width = unit(1, "cm"),
            legend.key.justification = "center",
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            panel.border = element_rect(colour = "black", 
                                        fill = NA, linewidth = 1))
  }
  return(p)
}
```

## Snow Goose {.unnumbered}

```{r SNGOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Snow goose (*Anser caerulescens*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")  
```

```{r SNGOfig, echo=FALSE, warning=FALSE}
#| label: fig-SNGO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line). 
 
spp = "SNGO"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Snow Goose {.unnumbered}

```{r SNGO, include=TRUE, echo=FALSE}
#| label: tbl-SNGO
cap <- paste0("Snow goose (*Anser caerulescens*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SNGO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Snow Goose {.unnumbered}

```{r SNGOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of snow geese (*Anser caerulescens*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SNGOmap, echo=FALSE, warning=FALSE}
#| label: fig-SNGOmap 
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Greater White-fronted Goose {.unnumbered}

```{r GWFGcap, echo=FALSE, warning=FALSE}
 cap = paste0("Greater white-fronted goose (*Anser albifrons*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r GWFGfig, echo=FALSE, warning=FALSE}
#| label: fig-GWFG
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "GWFG"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))
```

{{< pagebreak >}}

## Greater White-fronted Goose {.unnumbered}

```{r GWFG, include=TRUE, echo=FALSE}
#| label: tbl-GWFG
cap <- paste0("Greater white-fronted goose (*Anser albifrons*) indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(GWFG, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  #add_header adds a to tiered column header 'top' (e.g., instead of just 'Breeding Birds', the column header reads 'Indicated Breeding Birds'. This is not necessary for non-indicated species)
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Greater White-fronted Goose {.unnumbered}

```{r GWFGmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of greater white-fronted geese (*Anser albifrons*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r GWFGmap, echo=FALSE, warning=FALSE}
#| label: fig-GWFGmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r BRANcap, echo=FALSE, warning=FALSE}
 cap = paste0("Brant (*Branta bernicla nigricans*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r BRANfig, echo=FALSE, warning=FALSE}
#| label: fig-BRAN
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "BRAN"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))
```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r BRAN, include=TRUE, echo=FALSE}
#| label: tbl-BRAN
cap <- paste0("Brant (*Branta bernicla nigricans*) indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(BRAN, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r BRANmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of brant (*Branta bernicla nigricans*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r BRANmap, echo=FALSE, warning=FALSE}
#| label: fig-BRANmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Cackling/Canada Goose {.unnumbered}

```{r CCGOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Cackling/Canada goose (*Branta hutchinsii* and *B. canadensis*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r CCGOfig, echo=FALSE, warning=FALSE}
#| label: fig-CCGO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "CCGO"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))
```

{{< pagebreak >}}

## Cackling/Canada Goose {.unnumbered}

```{r CCGO, include=TRUE, echo=FALSE}
#| label: tbl-CCGO
cap <- paste0("Cackling/Canada goose (*Branta hutchinsii* and *B. canadensis*) indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(CCGO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Cackling/Canada Goose {.unnumbered}

```{r CCGOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of cackling/Canada geese (*Branta hutchinsii* and *B. canadensis*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r CCGOmap, echo=FALSE, warning=FALSE}
#| label: fig-CCGOmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Tundra Swan {.unnumbered}

```{r SWANcap, echo=FALSE, warning=FALSE}
 cap = paste0("Tundra swan (*Cygnus columbianus*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")  
```

```{r SWANfig, echo=FALSE, warning=FALSE}
#| label: fig-SWAN
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SWAN"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Tundra Swan {.unnumbered}

```{r SWAN, include=TRUE, echo=FALSE}
#| label: tbl-SWAN
cap <- paste0("Tundra swan (*Cygnus columbianus*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SWAN, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Tundra Swan {.unnumbered}

```{r SWANmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of tundra swans (*Cygnus columbianus*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SWANmap, echo=FALSE, warning=FALSE}
#| label: fig-SWANmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Northern Shoveler {.unnumbered}

```{r NSHOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Northern shoveler (*Spatula clypeata*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r NSHOfig, echo=FALSE, warning=FALSE}
#| label: fig-NSHO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "NSHO"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Northern Shoveler {.unnumbered}

```{r NSHO, include=TRUE, echo=FALSE}
#| label: tbl-NSHO
cap <- paste0("Northern shoveler (*Spatula clypeata*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(NSHO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Northern Shoveler {.unnumbered}

```{r NSHOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of northern shovelers (*Spatula clypeata*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r NSHOmap, echo=FALSE, warning=FALSE}
#| label: fig-NSHOmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## American Wigeon {.unnumbered}

```{r AMWIcap, echo=FALSE, warning=FALSE}
 cap = paste0("American wigeon (*Mareca americana*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r AMWIfig, echo=FALSE, warning=FALSE}
#| label: fig-AMWI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "AMWI"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## American Wigeon {.unnumbered}

```{r AMWI, include=TRUE, echo=FALSE}
#| label: tbl-AMWI
cap <- paste0("American wigeon (*Mareca americana*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(AMWI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## American Wigeon {.unnumbered}

```{r AMWImcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of American wigeon (*Mareca americana*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r AMWImap, echo=FALSE, warning=FALSE}
#| label: fig-AMWImap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r MALLcap, echo=FALSE, warning=FALSE}
 cap = paste0("Mallard (*Anas platyrhynchos*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r MALLfig, echo=FALSE, warning=FALSE}
#| label: fig-MALL
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "MALL"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r MALL, include=TRUE, echo=FALSE}
#| label: tbl-MALL
cap <- paste0("Mallard (*Anas platyrhynchos*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(MALL, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r MALLmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of mallards (*Anas platyrhynchos*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r MALLmap, echo=FALSE, warning=FALSE}
#| label: fig-MALLmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Northern Pintail {.unnumbered}

```{r NOPIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Northern pintail (*Anas acuta*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r NOPIfig, echo=FALSE, warning=FALSE}
#| label: fig-NOPI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "NOPI"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Northern Pintail {.unnumbered}

```{r NOPI, include=TRUE, echo=FALSE}
#| label: tbl-NOPI
cap <- paste0("Northern pintail (*Anas acuta*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(NOPI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Northern Pintail {.unnumbered}

```{r NOPImcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of northern pintails (*Anas acuta*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r NOPImap, echo=FALSE, warning=FALSE}
#| label: fig-NOPImap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## American Green-winged Teal {.unnumbered}

```{r GWTEcap, echo=FALSE, warning=FALSE}
 cap = paste0("American green-winged teal (*Anas crecca*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r GWTEfig, echo=FALSE, warning=FALSE}
#| label: fig-GWTE
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "GWTE"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## American Green-winged Teal {.unnumbered}

```{r GWTE, include=TRUE, echo=FALSE}
#| label: tbl-GWTE
cap <- paste0("American green-winged teal (*Anas crecca*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(GWTE, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "ccccc", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## American Green-winged Teal {.unnumbered}

```{r GWTEmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of American green-winged teal (*Anas crecca*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r GWTEmap, echo=FALSE, warning=FALSE}
#| label: fig-GWTEmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Scaup species {.unnumbered}

```{r UNSCcap, echo=FALSE, warning=FALSE}
 cap = paste0("Scaup species (*Aythya marila* and *A. affinis*) indices of breeding birds (grey circles; singles + [2 x pairs] + flocked drakes <5) and total birds (black circles; singles + [2 x pairs] + flocked drakes <5 + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r UNSCfig, echo=FALSE, warning=FALSE}
#| label: fig-UNSC
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "UNSC"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Scaup species {.unnumbered}

```{r UNSC, include=TRUE, echo=FALSE}
#| label: tbl-UNSC
cap <- paste0("Scaup species (*Aythya marila* and *A. affinis*) indices of breeding birds (singles + [2 x pairs] + flocked drakes <5) and total birds (singles + [2 x pairs] + flocked drakes <5 + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(UNSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Scaup species {.unnumbered}

```{r UNSCmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of scaup species (*Aythya marila* and *A. affinis*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r UNSCmap, echo=FALSE, warning=FALSE}
#| label: fig-UNSCmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Steller's Eider {.unnumbered}

```{r STEIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Steller’s eider (*Polysticta stelleri*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r STEIfig, echo=FALSE, warning=FALSE}
#| label: fig-STEI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "STEI"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Steller's Eider {.unnumbered}

```{r STEI, include=TRUE, echo=FALSE}
#| label: tbl-STEI
cap <- paste0("Steller’s eider (*Polysticta stelleri*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(STEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Steller's Eider {.unnumbered}

```{r STEImcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of Steller’s eiders (*Polysticta stelleri*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r STEImap, echo=FALSE, warning=FALSE}
#| label: fig-STEImap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Spectacled Eider {.unnumbered}

```{r SPEIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Spectacled eider (*Somateria fischeri*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SPEIfig, echo=FALSE, warning=FALSE}
#| label: fig-SPEI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SPEI"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Spectacled Eider {.unnumbered}

```{r SPEI, include=TRUE, echo=FALSE}
#| label: tbl-SPEI
cap <- paste0("Spectacled eider (*Somateria fischeri*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SPEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Spectacled Eider {.unnumbered}

```{r SPEImcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of spectacled eiders (*Somateria fischeri*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SPEImap, echo=FALSE, warning=FALSE}
#| label: fig-SPEImap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## King Eider {.unnumbered}

```{r KIEIcap, echo=FALSE, warning=FALSE}
 cap = paste0("King eider (*Somateria spectabilis*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r KIEIfig, echo=FALSE, warning=FALSE}
#| label: fig-KIEI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "KIEI"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## King Eider {.unnumbered}

```{r KIEI, include=TRUE, echo=FALSE}
#| label: tbl-KIEI
cap <- paste0("King eider (*Somateria spectabilis*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(KIEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## King Eider {.unnumbered}

```{r KIEImcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of king eiders (*Somateria spectabilis*) geese along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r KIEImap, echo=FALSE, warning=FALSE}
#| label: fig-KIEImap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Common Eider {.unnumbered}

```{r COEIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Common eider (*Somateria mollissima*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r COEIfig, echo=FALSE, warning=FALSE}
#| label: fig-COEI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "COEI"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Common Eider {.unnumbered}

```{r COEI, include=TRUE, echo=FALSE}
#| label: tbl-COEI
cap <- paste0("Common eider (*Somateria mollissima*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(COEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Common Eider {.unnumbered}

```{r COEImcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of common eiders (*Somateria mollissima*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r COEImap, echo=FALSE, warning=FALSE}
#| label: fig-COEImap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species. 

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Surf Scoter {.unnumbered}

```{r SUSCcap, echo=FALSE, warning=FALSE}
 cap = paste0("Surf scoter (*Melanitta perspicillata*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SUSCfig, echo=FALSE, warning=FALSE}
#| label: fig-SUSC
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SUSC"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Surf Scoter {.unnumbered}

```{r SUSC, include=TRUE, echo=FALSE}
#| label: tbl-SUSC
cap <- paste0("Surf scoter (*Melanitta perspicillata*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SUSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Surf Scoter {.unnumbered}

```{r SUSCmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of surf scoters (*Melanitta perspicillata*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SUSCmap, echo=FALSE, warning=FALSE}
#| label: fig-SUSCmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## White-winged Scoter {.unnumbered}

```{r WWSCcap, echo=FALSE, warning=FALSE}
 cap = paste0("White-winged scoter (*Melanitta deglandi*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r WWSCfig, echo=FALSE, warning=FALSE}
#| label: fig-WWSC
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "WWSC"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## White-winged Scoter {.unnumbered}

```{r WWSC, include=TRUE, echo=FALSE}
#| label: tbl-WWSC
cap <- paste0("White-winged scoter (*Melanitta deglandi*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(WWSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## White-winged Scoter {.unnumbered}

```{r WWSCmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of white-winged scoters (*Melanitta deglandi*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r WWSCmap, echo=FALSE, warning=FALSE}
#| label: fig-WWSCmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Black Scoter {.unnumbered}

```{r BLSCcap, echo=FALSE, warning=FALSE}
 cap = paste0("Black scoter (*Melanitta americana*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r BLSCfig, echo=FALSE, warning=FALSE}
#| label: fig-BLSC
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "BLSC"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Black Scoter {.unnumbered}

```{r BLSC, include=TRUE, echo=FALSE}
#| label: tbl-BLSC
cap <- paste0("Black scoter (*Melanitta americana*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(BLSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Black Scoter {.unnumbered}

```{r BLSCmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of black scoters (*Melanitta americana*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r BLSCmap, echo=FALSE, warning=FALSE}
#| label: fig-BLSCmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Long-tailed Duck {.unnumbered}

```{r LTDUcap, echo=FALSE, warning=FALSE}
 cap = paste0("Long-tailed duck (*Clangula hyemalis*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r LTDUfig, echo=FALSE, warning=FALSE}
#| label: fig-LTDU
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "LTDU"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Long-tailed Duck {.unnumbered}

```{r LTDU, include=TRUE, echo=FALSE}
#| label: tbl-LTDU
cap <- paste0("Long-tailed duck (*Clangula hyemalis*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(LTDU, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Long-tailed Duck {.unnumbered}

```{r LTDUmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of long-tailed ducks (*Clangula hyemalis*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r LTDUmap, echo=FALSE, warning=FALSE}
#| label: fig-LTDUmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Red-breasted Merganser {.unnumbered}

```{r RBMEcap, echo=FALSE, warning=FALSE}
 cap = paste0("Red-breasted merganser (*Mergus serrator*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r RBMEfig, echo=FALSE, warning=FALSE}
#| label: fig-RBME
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "RBME"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Red-breasted Merganser {.unnumbered}

```{r RBME, include=TRUE, echo=FALSE}
#| label: tbl-RBME
cap <- paste0("Red-breasted merganser (*Mergus serrator*) indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(RBME, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Red-breasted Merganser {.unnumbered}

```{r RBMEmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of red-breasted mergansers (*Mergus serrator*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r RBMEmap, echo=FALSE, warning=FALSE}
#| label: fig-RBMEmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Red-necked Grebe {.unnumbered}

```{r RNGRcap, echo=FALSE, warning=FALSE}
 cap = paste0("Red-necked grebe (*Podiceps grisegena*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r RNGRfig, echo=FALSE, warning=FALSE}
#| label: fig-RNGR
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "RNGR"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Red-necked Grebe {.unnumbered}

```{r RNGR, include=TRUE, echo=FALSE}
#| label: tbl-RNGR
cap <- paste0("Red-necked grebe (*Podiceps grisegena*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(RNGR, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Red-necked Grebe {.unnumbered}

```{r RNGRmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of red-necked grebes (*Podiceps grisegena*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r RNGRmap, echo=FALSE, warning=FALSE}
#| label: fig-RNGRmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Sandhill Crane {.unnumbered}

```{r SACRcap, echo=FALSE, warning=FALSE}
 cap = paste0("Sandhill crane (*Antigone canadensis*) indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SACRfig, echo=FALSE, warning=FALSE}
#| label: fig-SACR
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SACR"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se)  |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Sandhill Crane {.unnumbered}

```{r SACR, include=TRUE, echo=FALSE}
#| label: tbl-SACR
cap <- paste0("Sandhill crane (*Antigone canadensis*) indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SACR, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Sandhill Crane {.unnumbered}

```{r SACRmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of sandhill cranes (*Antigone canadensis*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SACRmap, echo=FALSE, warning=FALSE}
#| label: fig-SACRmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r JAEGcap, echo=FALSE, warning=FALSE}
 cap = paste0("Jaeger species (*Stercorarius sp.*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r JAEGfig, echo=FALSE, warning=FALSE}
#| label: fig-JAEG
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "JAEG"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r JAEG, include=TRUE, echo=FALSE}
#| label: tbl-JAEG
cap <- paste0("Jaeger species (*Stercorarius sp.*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(JAEG, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r JAEGmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of jaeger species (*Stercorarius sp.*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r JAEGmap, echo=FALSE, warning=FALSE}
#| label: fig-JAEGmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Sabine's Gull {.unnumbered}

```{r SAGUcap, echo=FALSE, warning=FALSE}
 cap = paste0("Sabine's gull (*Xema sabini*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SAGUfig, echo=FALSE, warning=FALSE}
#| label: fig-SAGU
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SAGU"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Sabine's Gull {.unnumbered}

```{r SAGU, include=TRUE, echo=FALSE}
#| label: tbl-SAGU
cap <- paste0("Sabine's gull (*Xema sabini*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SAGU, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Sabine's Gull {.unnumbered}

```{r SAGUmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of Sabine's gulls (*Xema sabini*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SAGUmap, echo=FALSE, warning=FALSE}
#| label: fig-SAGUmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Glaucous Gull {.unnumbered}

```{r GLGUcap, echo=FALSE, warning=FALSE}
 cap = paste0("Glaucous gull (*Larus hyperboreus*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r GLGUfig, echo=FALSE, warning=FALSE}
#| label: fig-GLGU
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "GLGU"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Glaucous Gull {.unnumbered}

```{r GLGU, include=TRUE, echo=FALSE}
#| label: tbl-GLGU
cap <- paste0("Glaucous gull (*Larus hyperboreus*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(GLGU, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Glaucous Gull {.unnumbered}

```{r GLGUmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of glaucous gulls (*Larus hyperboreus*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r GLGUmap, echo=FALSE, warning=FALSE}
#| label: fig-GLGUmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Arctic Tern {.unnumbered}

```{r ARTEcap, echo=FALSE, warning=FALSE}
 cap = paste0("Arctic tern (*Sterna paradisaea*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r ARTEfig, echo=FALSE, warning=FALSE}
#| label: fig-ARTE
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "ARTE"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Arctic Tern {.unnumbered}

```{r ARTE, include=TRUE, echo=FALSE}
#| label: tbl-ARTE
cap <- paste0("Arctic tern (*Sterna paradisaea*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(ARTE, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Arctic Tern {.unnumbered}

```{r ARTEmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of Arctic terns (*Sterna paradisaea*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r ARTEmap, echo=FALSE, warning=FALSE}
#| label: fig-ARTEmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Red-throated Loon {.unnumbered}

```{r RTLOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Red-throated loon (*Gavia stellata*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r RTLOfig, echo=FALSE, warning=FALSE}
#| label: fig-RTLO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "RTLO"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Red-throated Loon {.unnumbered}

```{r RTLO, include=TRUE, echo=FALSE}
#| label: tbl-RTLO
cap <- paste0("Red-throated loon (*Gavia stellata*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(RTLO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Red-throated Loon {.unnumbered}

```{r RTLOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of red-throated loons (*Gavia stellata*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r RTLOmap, echo=FALSE, warning=FALSE}
#| label: fig-RTLOmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Pacific Loon {.unnumbered}

```{r PALOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Pacific loon (*Gavia pacifica*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r PALOfig, echo=FALSE, warning=FALSE}
#| label: fig-PALO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "PALO"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Pacific Loon {.unnumbered}

```{r PALO, include=TRUE, echo=FALSE}
#| label: tbl-PALO
cap <- paste0("Pacific loon (*Gavia pacifica*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(PALO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Pacific Loon {.unnumbered}

```{r PALOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of Pacific loons (*Gavia pacifica*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r PALOmap, echo=FALSE, warning=FALSE}
#| label: fig-PALOmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Yellow-billed Loon {.unnumbered}

```{r YBLOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Yellow-billed loon (*Gavia adamsii*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r YBLOfig, echo=FALSE, warning=FALSE}
#| label: fig-YBLO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "YBLO"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Yellow-billed Loon {.unnumbered}

```{r YBLO, include=TRUE, echo=FALSE}
#| label: tbl-YBLO
cap <- paste0("Yellow-billed loon (*Gavia adamsii*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(YBLO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Yellow-billed Loon {.unnumbered}

```{r YBLOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of yellow-billed loons (*Gavia adamsii*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r YBLOmap, echo=FALSE, warning=FALSE}
#| label: fig-YBLOmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Golden Eagle {.unnumbered}

```{r GOEAcap, echo=FALSE, warning=FALSE}
 cap = paste0("Golden eagle (*Aquila chrysaetos*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r GOEAfig, echo=FALSE, warning=FALSE}
#| label: fig-GOEA
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "GOEA"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Golden Eagle {.unnumbered}

```{r GOEA, include=TRUE, echo=FALSE}
#| label: tbl-GOEA
cap <- paste0("Golden eagle (*Aquila chrysaetos*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(GOEA, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Golden Eagle {.unnumbered}

```{r GOEAmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of golden eagles (*Aquila chrysaetos*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r GOEAmap, echo=FALSE, warning=FALSE}
#| label: fig-GOEAmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Short-eared Owl {.unnumbered}

```{r SEOWcap, echo=FALSE, warning=FALSE}
 cap = paste0("Short-eared owl (*Asio flammeus*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SEOWfig, echo=FALSE, warning=FALSE}
#| label: fig-SEOW
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SEOW"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Short-eared Owl {.unnumbered}

```{r SEOW, include=TRUE, echo=FALSE}
#| label: tbl-SEOW
cap <- paste0("Short-eared owl (*Asio flammeus*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SEOW, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Short-eared Owl {.unnumbered}

```{r SEOWmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of short-eared owls (*Asio flammeus*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SEOWmap, echo=FALSE, warning=FALSE}
#| label: fig-SEOWmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Snowy Owl {.unnumbered}

```{r SNOWcap, echo=FALSE, warning=FALSE}
 cap = paste0("Snowy owl (*Bubo scandiacus*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SNOWfig, echo=FALSE, warning=FALSE}
#| label: fig-SNOW
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "SNOW"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Snowy Owl {.unnumbered}

```{r SNOW, include=TRUE, echo=FALSE}
#| label: tbl-SNOW
cap <- paste0("Snowy owl (*Bubo scandiacus*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(SNOW, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Snowy Owl {.unnumbered}

```{r SNOWmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of snowy owls (*Bubo scandiacus*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r SNOWmap, echo=FALSE, warning=FALSE}
#| label: fig-SNOWmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

## Common Raven {.unnumbered}

```{r CORAcap, echo=FALSE, warning=FALSE}
 cap = paste0("Common raven (*Corvus corax*) indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(ACPHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007–",max(ACPHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r CORAfig, echo=FALSE, warning=FALSE}
#| label: fig-CORA
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2025 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "CORA"
df <- AKaerial::ACPHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2025, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Common Raven {.unnumbered}

```{r CORA, include=TRUE, echo=FALSE}
#| label: tbl-CORA
cap <- paste0("Common raven (*Corvus corax*) indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, 2007-",max(ACPHistoric$combined$Year),".")

kable(CORA, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Common Raven {.unnumbered}

```{r CORAmcap, echo=FALSE, warning=FALSE}
 cap = paste0("(A) Observations of common ravens (*Corvus corax*) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) from the Arctic Coastal Plain Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(ACPHistoric$combined$Year),"). (B) Expected density (2007–",max(ACPHistoric$combined$Year),") of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species." )
```

```{r CORAmap, echo=FALSE, warning=FALSE}
#| label: fig-CORAmap
#| fig-width: 7
#| fig-height: 7
#| fig-cap: !expr cap
#| fig-alt: In (A) is a map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and Teshekpuk high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year. In (B) is a map of expected density (all survey years) of indicated breeding pairs (number of single birds + number of bird pairs) on the natural log scale across the survey area predicted from a generalized additive model after removing observer, year, and spatio-temporal effects. The scale bar shows density of indicated breeding pairs per square kilometer from the minimum to maximum value for this species.  

p1 <- AKaerial::ObsMap(area="ACP", species=spp) + coord_sf(crs = 3338)

p2 <- plot_raster(x = paste0("data/gam/ACP-", spp, "-2026-04-06.tif"), 
                  what = "gg", log = TRUE)

p1/p2 + plot_annotation(tag_levels = 'A')
```

{{< pagebreak >}}

