
# APPENDICES {.unnumbered}

```{r}
#library(AKaerial)
#library(tidyverse)
#library(kableExtra)

#load("data/YKD_Report.RData")

cap2 = paste0(" indices of indicated breeding birds (grey circles; 2 x [singles + 
              pairs]) and indicated total birds (black circles; 2 x [singles + pairs] 
              + birds in flocks) with 95% confidence intervals (vertical bars from 
              circles), as well as the long-term trend in indicated total birds (black 
              line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible 
              intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial 
              Breeding Population Survey, Alaska, 1985–",max(YKGHistoric$combined$Year),
              ". Years with counts of zero were not included in trend estimates 
              (see Methods for details).")

cap3 <- paste0(" indices of breeding birds (singles + [2 x pairs]) and total birds 
               (singles + [2 x pairs] + birds in flocks) with standard errors (SE) 
               from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 
               1985-",max(YKGHistoric$combined$Year),".")

cap4 = paste0(" along transects within four physiographic-based strata (high, medium, 
              low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding 
              Population Survey, Alaska, during the most recent survey year (",
              max(YKGHistoric$combined$Year),")." )

alttext = paste0("A graph of the average index estimate between observers (vertical bars 
from circles), the posterior mean population growth trajectory 1985-", 
max(YKGHistoric$combined$Year)," (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).")

altmap = paste0("A map of species-specific observation locations (black dots) along 
transects within four physiographic-based strata (high, medium, low, and very high) 
of the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the 
most recent survey year (", max(YKGHistoric$combined$Year),")." )
```

## Emperor goose {.unnumbered}

```{r}
#| label: fig-EMGO
#| fig-cap: !expr paste( "Emperor goose", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "EMGO", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Emperor goose data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-EMGO

kable(EMGO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Emperor goose", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Emperor goose map {.unnumbered}

```{r}
#| label: fig-EMGOmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Emperor goose", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "EMGO", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Greater white-fronted goose {.unnumbered}

```{r}
#| label: fig-GWFG
#| fig-cap: !expr paste( "Greater white-fronted goose", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "GWFG", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Greater white-fronted goose data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-GWFG

kable(GWFG, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Greater white-fronted goose", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Greater white-fronted goose map {.unnumbered}

```{r}
#| label: fig-GWFGmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Greater white-fronted goose", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "GWFG", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r}
#| label: fig-BRAN
#| fig-cap: !expr paste( "Brant", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "BRAN", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Brant data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-BRAN

kable(BRAN, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Brant", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Brant map {.unnumbered}

```{r}
#| label: fig-BRANmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Brant", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "BRAN", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Cackling Canada goose {.unnumbered}

```{r}
#| label: fig-CCGO
#| fig-cap: !expr paste( "Cackling Canada goose", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "CCGO", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Cackling Canada goose data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-CCGO

kable(CCGO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Cackling Canada goose", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Cackling Canada goose map {.unnumbered}

```{r}
#| label: fig-CCGOmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Cackling Canada goose", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "CCGO", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Taverner's Canada goose {.unnumbered}

```{r}
#| label: fig-TAVS
#| fig-cap: !expr paste( "Taverner's Canada goose", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "TAVS", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Taverner's Canada goose data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-TAVS

kable(TAVS, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Taverner's Canada goose", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Taverner's Canada goose map {.unnumbered}

```{r}
#| label: fig-TAVSmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Taverner's Canada goose", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "TAVS", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Tundra swan {.unnumbered}

```{r}
#| label: fig-SWAN
#| fig-cap: !expr paste( "Tundra swan", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "SWAN", index = "total", index2 = "sing1pair2" )
```

{{< pagebreak >}}

### Tundra swan data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-SWAN

kable(SWAN, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Tundra swan", cap3), align = "lcccc", 
  col.names = c("Year", "Total Birds", "SE", "Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Tundra swan map {.unnumbered}

```{r}
#| label: fig-SWANmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Tundra swan", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "SWAN", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## American wigeon {.unnumbered}

```{r}
#| label: fig-AMWI
#| fig-cap: !expr paste( "American wigeon", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "AMWI", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### American wigeon data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-AMWI

kable(AMWI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "American wigeon", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### American wigeon map {.unnumbered}

```{r}
#| label: fig-AMWImap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "American wigeon", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "AMWI", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r}
#| label: fig-MALL
#| fig-cap: !expr paste( "Mallard", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "MALL", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Mallard data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-MALL

kable(MALL, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Mallard", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Mallard map {.unnumbered}

```{r}
#| label: fig-MALLmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Mallard", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "MALL", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Northern pintail {.unnumbered}

```{r}
#| label: fig-NOPI
#| fig-cap: !expr paste( "Northern pintail", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "NOPI", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Northern pintail data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-NOPI

kable(NOPI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Northern pintail", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Northern pintail map {.unnumbered}

```{r}
#| label: fig-NOPImap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Northern pintail", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "NOPI", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Green-winged teal {.unnumbered}

```{r}
#| label: fig-GWTE
#| fig-cap: !expr paste( "Green-winged teal", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "GWTE", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Green-winged teal data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-GWTE

kable(GWTE, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Green-winged teal", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Green-winged teal map {.unnumbered}

```{r}
#| label: fig-GWTEmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Green-winged teal", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "GWTE", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Scaup species {.unnumbered}

```{r}
#| label: fig-UNSC
#| fig-cap: !expr paste( "Scaup species", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "UNSC", index = "total", index2 = "sing1pair2" )
```

{{< pagebreak >}}

### Scaup species data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-UNSC

kable(UNSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Scaup species", cap3), align = "lcccc", 
  col.names = c("Year", "Total Birds", "SE", "Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Scaup species map {.unnumbered}

```{r}
#| label: fig-UNSCmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Scaup species", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "UNSC", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Spectacled eider {.unnumbered}

```{r}
#| label: fig-SPEI
#| fig-cap: !expr paste( "Spectacled eider", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "SPEI", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Spectacled eider data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-SPEI

kable(SPEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Spectacled eider", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Spectacled eider map {.unnumbered}

```{r}
#| label: fig-SPEImap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Spectacled eider", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "SPEI", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Common eider {.unnumbered}

```{r}
#| label: fig-COEI
#| fig-cap: !expr paste( "Common eider", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "COEI", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Common eider data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-COEI

kable(COEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Common eider", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Common eider map {.unnumbered}

```{r}
#| label: fig-COEImap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Common eider", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "COEI", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Black scoter {.unnumbered}

```{r}
#| label: fig-BLSC
#| fig-cap: !expr paste( "Black scoter", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "BLSC", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Black scoter data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-BLSC

kable(BLSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Black scoter", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Black scoter map {.unnumbered}

```{r}
#| label: fig-BLSCmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Black scoter", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "BLSC", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Long-tailed duck {.unnumbered}

```{r}
#| label: fig-LTDU
#| fig-cap: !expr paste( "Long-tailed duck", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "LTDU", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Long-tailed duck data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-LTDU

kable(LTDU, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Long-tailed duck", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Long-tailed duck map {.unnumbered}

```{r}
#| label: fig-LTDUmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Long-tailed duck", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "LTDU", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Red-breasted merganser {.unnumbered}

```{r}
#| label: fig-RBME
#| fig-cap: !expr paste( "Red-breasted merganser", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "RBME", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Red-breasted merganser data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-RBME

kable(RBME, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Red-breasted merganser", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Red-breasted merganser map {.unnumbered}

```{r}
#| label: fig-RBMEmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Red-breasted merganser", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "RBME", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Sandhill crane {.unnumbered}

```{r}
#| label: fig-SACR
#| fig-cap: !expr paste( "Sandhill crane", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "SACR", index = "itotal", index2 = "ibb" )
```

{{< pagebreak >}}

### Sandhill crane data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-SACR

kable(SACR, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Sandhill crane", cap3), align = "lcccc", 
  col.names = c("Year", "Indicated Total Birds", "SE", "Indicated Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Sandhill crane map {.unnumbered}

```{r}
#| label: fig-SACRmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Sandhill crane", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKG", species = "SACR", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r}
#| label: fig-JAEG
#| fig-cap: !expr paste( "Jaeger species", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "JAEG", index = "total", index2 = "sing1pair2" )
```

{{< pagebreak >}}

### Jaeger species data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-JAEG

kable(JAEG, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Jaeger species", cap3), align = "lcccc", 
  col.names = c("Year", "Total Birds", "SE", "Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Jaeger species map {.unnumbered}

```{r}
#| label: fig-JAEGmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Jaeger species", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "JAEG", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Red-throated loon {.unnumbered}

```{r}
#| label: fig-RTLO
#| fig-cap: !expr paste( "Red-throated loon", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "RTLO", index = "total", index2 = "sing1pair2" )
```

{{< pagebreak >}}

### Red-throated loon data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-RTLO

kable(RTLO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Red-throated loon", cap3), align = "lcccc", 
  col.names = c("Year", "Total Birds", "SE", "Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Red-throated loon map {.unnumbered}

```{r}
#| label: fig-RTLOmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Red-throated loon", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "RTLO", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}

## Pacific loon {.unnumbered}

```{r}
#| label: fig-PALO
#| fig-cap: !expr paste( "Pacific loon", cap2)
#| fig-alt: !expr alttext

sppfig(spp = "PALO", index = "total", index2 = "sing1pair2" )
```

{{< pagebreak >}}

### Pacific loon data table {.unnumbered}

```{r}
#| include: true
#| label: tbl-PALO

kable(PALO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), 
  caption = paste0( "Pacific loon", cap3), align = "lcccc", 
  col.names = c("Year", "Total Birds", "SE", "Breeding Birds", "SE")) %>%
  kable_styling(font_size=10, position="center") %>%
  row_spec(row = 0, bold = TRUE)

```
{{< pagebreak >}}

### Pacific loon map {.unnumbered}

```{r}
#| label: fig-PALOmap
#| fig.height: 5.5
#| fig-cap: !expr paste0( "Pacific loon", cap4)
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

ObsMap2(area = "YKD", species = "PALO", 
                 mirror = "data/mbmwa_001_YKD_Aerial_Survey/data/")  + 
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

{{< pagebreak >}}


