# APPENDICES {.unnumbered}

## Emperor Goose {.unnumbered}

```{r cap2}
 cap2 = paste0(" indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKGHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r EMGOfig}
#| label: fig-SNGO
#| fig-cap: !expr paste0("Emperor goose",cap2)
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 1985-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line). 

sppfig(spp = "EMGO")
```

{{< pagebreak >}}

### Emperor Goose Data Table {.unnumbered}

```{r EMGO, include=TRUE, echo=FALSE}
#| label: tbl-EMGO
cap <- paste0("Emperor goose indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKGHistoric$combined$Year),".")

kable(EMGO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

### Emperor Goose Map {.unnumbered}

```{r mcap}
 cap3 = paste0("Observations of emperor geese along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKGHistoric$combined$Year),")." )
```

```{r EMGOmap, echo=FALSE, warning=FALSE}
#| label: fig-EMGOmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKG", species="EMGO")

```

{{< pagebreak >}}

## Greater White-fronted Goose {.unnumbered}

```{r GWFGcap, echo=FALSE, warning=FALSE}
 cap = paste0("Greater white-fronted goose indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKGHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r GWFGfig, echo=FALSE, warning=FALSE}
#| label: fig-GWFG
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line).
 
spp = "GWFG"
df <- AKaerial::YKGHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))
```

{{< pagebreak >}}

## Greater White-fronted Goose {.unnumbered}

```{r GWFG, include=TRUE, echo=FALSE}
#| label: tbl-GWFG
cap <- paste0("Greater white-fronted goose indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKGHistoric$combined$Year),".")

kable(GWFG, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  #add_header adds a to tiered column header 'top' (e.g., instead of just 'Breeding Birds', the column header reads 'Indicated Breeding Birds'. This is not necessary for non-indicated species)
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Greater White-fronted Goose {.unnumbered}

```{r GWFGmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of greater white-fronted geese along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKGHistoric$combined$Year),")." )
```

```{r GWFGmap, echo=FALSE, warning=FALSE}
#| label: fig-GWFGmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKG", species="GWFG")

```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r BRANcap, echo=FALSE, warning=FALSE}
 cap = paste0("Brant indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKGHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r BRANfig, echo=FALSE, warning=FALSE}
#| label: fig-BRAN
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "BRAN"
df <- AKaerial::YKGHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))
```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r BRAN, include=TRUE, echo=FALSE}
#| label: tbl-BRAN
cap <- paste0("Brant indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKGHistoric$combined$Year),".")

kable(BRAN, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Brant {.unnumbered}

```{r BRANmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of brant along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKGHistoric$combined$Year),")." )
```

```{r BRANmap, echo=FALSE, warning=FALSE}
#| label: fig-BRANmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKG", species="BRAN")

```

{{< pagebreak >}}

## Cackling/Canada Goose {.unnumbered}

```{r CCGOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Cackling/Canada goose indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKGHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r CCGOfig, echo=FALSE, warning=FALSE}
#| label: fig-CCGO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "CCGO"
df <- AKaerial::YKGHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))
```

{{< pagebreak >}}

## Cackling/Canada Goose {.unnumbered}

```{r CCGO, include=TRUE, echo=FALSE}
#| label: tbl-CCGO
cap <- paste0("Cackling/Canada goose indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKGHistoric$combined$Year),".")

kable(CCGO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Cackling/Canada Goose {.unnumbered}

```{r CCGOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of cackling/Canada geese along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKGHistoric$combined$Year),")." )
```

```{r CCGOmap, echo=FALSE, warning=FALSE}
#| label: fig-CCGOmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKG", species="CCGO")

```

{{< pagebreak >}}

## Tundra Swan {.unnumbered}

```{r SWANcap, echo=FALSE, warning=FALSE}
 cap = paste0("Tundra swan indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKGHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")  
```

```{r SWANfig, echo=FALSE, warning=FALSE}
#| label: fig-SWAN
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "SWAN"
df <- AKaerial::YKGHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Tundra Swan {.unnumbered}

```{r SWAN, include=TRUE, echo=FALSE}
#| label: tbl-SWAN
cap <- paste0("Tundra swan indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKGHistoric$combined$Year),".")

kable(SWAN, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Tundra Swan {.unnumbered}

```{r SWANmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of tundra swans along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKGHistoric$combined$Year),")." )
```

```{r SWANmap, echo=FALSE, warning=FALSE}
#| label: fig-SWANmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKG", species="SWAN")

```

{{< pagebreak >}}

## Northern Shoveler {.unnumbered}

```{r NSHOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Northern shoveler indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r NSHOfig, echo=FALSE, warning=FALSE}
#| label: fig-NSHO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "NSHO"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Northern Shoveler {.unnumbered}

```{r NSHO, include=TRUE, echo=FALSE}
 
cap <- paste0("Northern shoveler indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(NSHO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Northern Shoveler {.unnumbered}

```{r NSHOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of northern shoveler along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r NSHOmap, echo=FALSE, warning=FALSE}
#| label: fig-NSHOmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="NSHO")

```

{{< pagebreak >}}

## American Wigeon {.unnumbered}

```{r AMWIcap, echo=FALSE, warning=FALSE}
 cap = paste0("American wigeon indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r AMWIfig, echo=FALSE, warning=FALSE}
#| label: fig-AMWI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "AMWI"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## American Wigeon {.unnumbered}

```{r AMWI, include=TRUE, echo=FALSE}
#| label: tbl-AMWI
cap <- paste0("American wigeon indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(AMWI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## American Wigeon {.unnumbered}

```{r AMWImcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of American wigeon along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r AMWImap, echo=FALSE, warning=FALSE}
#| label: fig-AMWImap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="AMWI")

```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r MALLcap, echo=FALSE, warning=FALSE}
 cap = paste0("Mallard indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r MALLfig, echo=FALSE, warning=FALSE}
#| label: fig-MALL
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "MALL"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r MALL, include=TRUE, echo=FALSE}
#| label: tbl-MALL
cap <- paste0("Mallard indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(MALL, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Mallard {.unnumbered}

```{r MALLmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of mallards along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r MALLmap, echo=FALSE, warning=FALSE}
#| label: fig-MALLmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="MALL")

```

{{< pagebreak >}}

## Northern Pintail {.unnumbered}

```{r NOPIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Northern pintail indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r NOPIfig, echo=FALSE, warning=FALSE}
#| label: fig-NOPI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "NOPI"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Northern Pintail {.unnumbered}

```{r NOPI, include=TRUE, echo=FALSE}
#| label: tbl-NOPI
cap <- paste0("Northern pintail indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(NOPI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Northern Pintail {.unnumbered}

```{r NOPImcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of northern pintails along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r NOPImap, echo=FALSE, warning=FALSE}
#| label: fig-NOPImap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="NOPI")

```

{{< pagebreak >}}

## American Green-winged Teal {.unnumbered}

```{r GWTEcap, echo=FALSE, warning=FALSE}
 cap = paste0("American green-winged teal indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r GWTEfig, echo=FALSE, warning=FALSE}
#| label: fig-GWTE
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "GWTE"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## American Green-winged Teal {.unnumbered}

```{r GWTE, include=TRUE, echo=FALSE}
#| label: tbl-GWTE
cap <- paste0("American green-winged teal indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(GWTE, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "ccccc", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## American Green-winged Teal {.unnumbered}

```{r GWTEmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of American green-winged teal along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r GWTEmap, echo=FALSE, warning=FALSE}
#| label: fig-GWTEmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="GWTE")

```












{{< pagebreak >}}



## Canvasback {.unnumbered}

```{r CANVcap, echo=FALSE, warning=FALSE}
 cap = paste0("Canvasback indices of breeding birds (grey circles; singles + [2 x pairs] + flocked drakes <5) and total birds (black circles; singles + [2 x pairs] + flocked drakes <5 + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r CANVfig, echo=FALSE, warning=FALSE}
#| label: fig-CANV
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "CANV"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Canvasback {.unnumbered}

```{r CANV, include=TRUE, echo=FALSE}
#| label: tbl-CANV
cap <- paste0("Canvasback indices of breeding birds (singles + [2 x pairs] + flocked drakes <5) and total birds (singles + [2 x pairs] + flocked drakes <5 + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(CANV, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Canvasback {.unnumbered}

```{r CANVmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of canvasback along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r CANVmap, echo=FALSE, warning=FALSE}
#| label: fig-CANVmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="CANV")

```

{{< pagebreak >}}


## Scaup species {.unnumbered}

```{r UNSCcap, echo=FALSE, warning=FALSE}
 cap = paste0("Scaup species indices of breeding birds (grey circles; singles + [2 x pairs] + flocked drakes <5) and total birds (black circles; singles + [2 x pairs] + flocked drakes <5 + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r UNSCfig, echo=FALSE, warning=FALSE}
#| label: fig-UNSC
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "UNSC"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Scaup species {.unnumbered}

```{r UNSC, include=TRUE, echo=FALSE}
#| label: tbl-UNSC
cap <- paste0("Scaup species indices of breeding birds (singles + [2 x pairs] + flocked drakes <5) and total birds (singles + [2 x pairs] + flocked drakes <5 + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(UNSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Scaup species {.unnumbered}

```{r UNSCmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of scaup species along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r UNSCmap, echo=FALSE, warning=FALSE}
#| label: fig-UNSCmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="UNSC")

```

{{< pagebreak >}}


## Spectacled Eider {.unnumbered}

```{r SPEIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Spectacled eider indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SPEIfig, echo=FALSE, warning=FALSE}
#| label: fig-SPEI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "SPEI"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Spectacled Eider {.unnumbered}

```{r SPEI, include=TRUE, echo=FALSE}
#| label: tbl-SPEI
cap <- paste0("Spectacled eider indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(SPEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Spectacled Eider {.unnumbered}

```{r SPEImcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of spectacled eiders along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r SPEImap, echo=FALSE, warning=FALSE}
#| label: fig-SPEImap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="SPEI")

```

{{< pagebreak >}}

## Common Eider {.unnumbered}

```{r COEIcap, echo=FALSE, warning=FALSE}
 cap = paste0("Common eider indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r COEIfig, echo=FALSE, warning=FALSE}
#| label: fig-COEI
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "COEI"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Common Eider {.unnumbered}

```{r COEI, include=TRUE, echo=FALSE}
#| label: tbl-COEI
cap <- paste0("Common eider indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(COEI, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Common Eider {.unnumbered}

```{r COEImcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of common eiders along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r COEImap, echo=FALSE, warning=FALSE}
#| label: fig-COEImap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="COEI")

```

{{< pagebreak >}}

## Black Scoter {.unnumbered}

```{r BLSCcap, echo=FALSE, warning=FALSE}
 cap = paste0("Black scoter indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r BLSCfig, echo=FALSE, warning=FALSE}
#| label: fig-BLSC
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "BLSC"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Black Scoter {.unnumbered}

```{r BLSC, include=TRUE, echo=FALSE}
#| label: tbl-BLSC
cap <- paste0("Black scoter indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(BLSC, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Black Scoter {.unnumbered}

```{r BLSCmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of black scoters along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r BLSCmap, echo=FALSE, warning=FALSE}
#| label: fig-BLSCmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="BLSC")

```

{{< pagebreak >}}

## Long-tailed Duck {.unnumbered}

```{r LTDUcap, echo=FALSE, warning=FALSE}
 cap = paste0("Long-tailed duck indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r LTDUfig, echo=FALSE, warning=FALSE}
#| label: fig-LTDU
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "LTDU"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Long-tailed Duck {.unnumbered}

```{r LTDU, include=TRUE, echo=FALSE}
#| label: tbl-LTDU
cap <- paste0("Long-tailed duck indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(LTDU, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Long-tailed Duck {.unnumbered}

```{r LTDUmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of long-tailed ducks along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r LTDUmap, echo=FALSE, warning=FALSE}
#| label: fig-LTDUmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="LTDU")

```

{{< pagebreak >}}

## Red-breasted Merganser {.unnumbered}

```{r RBMEcap, echo=FALSE, warning=FALSE}
 cap = paste0("Red-breasted merganser indices of indicated breeding birds (grey circles; 2 x [singles + pairs + flocked drakes <5]) and indicated total birds (black circles; 2 x [singles + pairs + flocked drakes <5] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r RBMEfig, echo=FALSE, warning=FALSE}
#| label: fig-RBME
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "RBME"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) + 
#scale_y_continuous discontinuous use of scientific notation to abbreviate large-numbered axis labels  
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank()) +
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Red-breasted Merganser {.unnumbered}

```{r RBME, include=TRUE, echo=FALSE}
#| label: tbl-RBME
cap <- paste0("Red-breasted merganser indices of indicated breeding birds (2 x [singles + pairs + flocked drakes <5]) and indicated total birds (2 x [singles + pairs + flocked drakes <5] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(RBME, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Red-breasted Merganser {.unnumbered}

```{r RBMEmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of red-breasted mergansers along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r RBMEmap, echo=FALSE, warning=FALSE}
#| label: fig-RBMEmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="RBME")

```

{{< pagebreak >}}


## Sandhill Crane {.unnumbered}

```{r SACRcap, echo=FALSE, warning=FALSE}
 cap = paste0("Sandhill crane indices of indicated breeding birds (grey circles; 2 x [singles + pairs]) and indicated total birds (black circles; 2 x [singles + pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in indicated total birds (black line; 2007–",max(YKGHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKGHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r SACRfig, echo=FALSE, warning=FALSE}
#| label: fig-SACR
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "SACR"
df <- AKaerial::YKGHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = itotal, se = itotal.se, index2 = ibb, se2 = ibb.se)  |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Indicated Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Indicated Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Indicated Total"= "black", "Indicated Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Sandhill Crane {.unnumbered}

```{r SACR, include=TRUE, echo=FALSE}
#| label: tbl-SACR
cap <- paste0("Sandhill crane indices of indicated breeding birds (2 x [singles + pairs]) and indicated total birds (2 x [singles + pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKGHistoric$combined$Year),".")

kable(SACR, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  add_header_above(bold = TRUE, c(" "=1, "Indicated"=1, " "=1, "Indicated" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Sandhill Crane {.unnumbered}

```{r SACRmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of sandhill cranes along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKGHistoric$combined$Year),")." )
```

```{r SACRmap, echo=FALSE, warning=FALSE}
#| label: fig-SACRmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKG", species="SACR")

```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r JAEGcap, echo=FALSE, warning=FALSE}
 cap = paste0("Jaeger species indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r JAEGfig, echo=FALSE, warning=FALSE}
#| label: fig-JAEG
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "JAEG"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r JAEG, include=TRUE, echo=FALSE}
#| label: tbl-JAEG
cap <- paste0("Jaeger species indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(JAEG, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Jaeger species {.unnumbered}

```{r JAEGmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of jaeger species along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r JAEGmap, echo=FALSE, warning=FALSE}
#| label: fig-JAEGmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="JAEG")

```

{{< pagebreak >}}

## Red-throated Loon {.unnumbered}

```{r RTLOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Red-throated loon indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r RTLOfig, echo=FALSE, warning=FALSE}
#| label: fig-RTLO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "RTLO"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Red-throated Loon {.unnumbered}

```{r RTLO, include=TRUE, echo=FALSE}
#| label: tbl-RTLO
cap <- paste0("Red-throated loon indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(RTLO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Red-throated Loon {.unnumbered}

```{r RTLOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of red-throated loons along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r RTLOmap, echo=FALSE, warning=FALSE}
#| label: fig-RTLOmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="RTLO")

```

{{< pagebreak >}}

## Pacific Loon {.unnumbered}

```{r PALOcap, echo=FALSE, warning=FALSE}
 cap = paste0("Pacific loon indices of breeding birds (grey circles; singles + [2 x pairs]) and total birds (black circles; singles + [2 x pairs] + birds in flocks) with 95% confidence intervals (vertical bars from circles), as well as the long-term trend in total birds (black line; 2007–",max(YKDHistoric$combined$Year),") and associated 95% credible intervals (grey band around line) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007–",max(YKDHistoric$combined$Year),". Years with counts of zero were not included in trend estimates (see Methods for details).")
```

```{r PALOfig, echo=FALSE, warning=FALSE}
#| label: fig-PALO
#| fig-cap: !expr cap
#| fig-alt: A graph of the average index estimate between observers (vertical bars from circles), the posterior mean population growth trajectory 2007-2024 (black line), and the approximate 95% confidence interval for the trajectory (grey band around line)..
 
spp = "PALO"
df <- AKaerial::YKDHistoric$combined |>
  filter(Species == spp) |>
  select(Year, index = total, se = total.se, index2 = sing1pair2, se2 = sing1pair2.se) |> 
  mutate(upper = index +1.96*se, lower = index - 1.96*se, upper2 = index2+1.96*se2, 
         lower2 = index2 - 1.96*se2)
df2 <- read_csv(file = paste0("data/plot_trends/data/",spp,"_bootfit.csv"))
ggplot(data = df2) + 
  geom_ribbon(aes(x=Year, ymin=pmax(0, lower), ymax = upper), fill = "lightgray", 
              alpha = 0.5) + 
  geom_line(aes(x=Year, y=mean), lwd = 1) + 
  geom_pointrange(data = df, aes(x=Year-0.1, y=index, ymin=pmax(0,lower), ymax = upper, 
                  col = "Total")) + 
  geom_pointrange(data = df, aes(x=Year+0.1, y=index2, ymin=pmax(0,lower2), ymax = upper2, 
                  col = "Breeding")) +
  labs(x = "Year", y = "Population Index") +
  scale_x_continuous(breaks = seq(2007, 2024, by = 2)) +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_manual(values=c("Total"= "black", "Breeding" = "darkgray")) +  
  theme_bw() + 
  theme(legend.position = "top", legend.title=element_blank(), legend.text=element_text(size=11)) +
  theme(text=element_text(family="Times")) +
  theme(axis.title=element_text(size=13,face="bold")) +
  theme(axis.text.x = element_text(size=11), axis.text.y = element_text(size=11))

```

{{< pagebreak >}}

## Pacific Loon {.unnumbered}

```{r PALO, include=TRUE, echo=FALSE}
#| label: tbl-PALO
cap <- paste0("Pacific loon indices of breeding birds (singles + [2 x pairs]) and total birds (singles + [2 x pairs] + birds in flocks) with standard errors (SE) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, 2007-",max(YKDHistoric$combined$Year),".")

kable(PALO, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, format.args = list(big.mark = ","), caption = cap, align = "lrrrr", col.names = c("Year", "Breeding Birds", "SE", "Total Birds", "SE")) %>%
  kable_styling(font_size=10, position="center")%>%
  row_spec(row = 0, bold = TRUE)

```

{{< pagebreak >}}

## Pacific Loon {.unnumbered}

```{r PALOmcap, echo=FALSE, warning=FALSE}
 cap = paste0("Observations of Pacific loons along transects within four physiographic-based strata (high, medium, low, and very high) from the Yukon-Kuskokwim Delta Aerial Breeding Population Survey, Alaska, during the most recent survey year (",max(YKDHistoric$combined$Year),")." )
```

```{r PALOmap, echo=FALSE, warning=FALSE}
#| label: fig-PALOmap
#| fig-cap: !expr cap
#| fig-alt: A map of species-specific observation locations (black dots) along transects within four physiographic-based strata (high, medium, low, and very high) of the Arctic Coastal Aerial Plain Breeding Population Survey, Alaska, during the most recent survey year (2024). 

AKaerial::ObsMap(area="YKD", species="PALO")

```


