---
title: "POPULATION INDICES, TRENDS, AND DISTRIBUTION OF BREEDING WATERBIRDS ON THE YUKON-KUSKOKWIM DELTA, ALASKA, 1985-2024"
affiliation:
  - id: fws
    name: U.S Fish and Wildlife Service
    department: Migratory Bird Management
    city: "Anchorage, AK 99503"
authors:
  - name: HEATHER M. WILSON
    email: heather_wilson@fws.gov
    orcid: 0009-0000-6387-0829
    affiliation:
      - ref: fws
  - name: MICHAEL A. SWAIM
    email: michael_swaim@fws.gov
    orcid: 0000-0002-9644-7127
    affiliation:
      - ref: fws
  - name: DAVID E. SAFINE
    email: david_safine@fws.gov
    orcid: 0009-0000-5648-0363
    affiliation:
      - ref: fws
  - name: CHARLES J. FROST
    email: charles_frost@fws.gov
    orcid: 0000-0002-6642-8371
    affiliation:
      - ref: fws
  - name: TAMARA K. ZELLER
    email: tamara_zeller@fws.gov
    affiliation:
      - ref: fws
  - name: ERIK E. OSNAS
    email: erik_osnas@fws.gov
    orcid: 0000-0001-9528-0866
    affiliation:
      - ref: fws
affiliations:
  - id: fws
    name: U.S Fish and Wildlife Service
    department: Migratory Bird Management
    city: "Anchorage, AK 99503"
date: ""
bibliography: "bibliography.bib"
bibliographystyle: apalike
toc: false
lof: false
lot: false
fig-cap-location: bottom
tbl-cap-location: top
format:
  arxiv-pdf:
    runninghead: "Yukon-Kuskokwim Delta Aerial Breeding Population Survey Report 1985-2024"
    number_sections: false
    keep-tex: true  
    fig-width: 8
    fig-height: 5
    include-in-header:
      text: |
        \setlength{\defaultaddspace}{0pt}
prefer-html: true
editor: source

embed-resources: true
linenumbers: true
execute:
  message: false
  warning: false
  echo: false
  
---

**Migratory Bird Management, U.S Fish and Wildlife Service, Anchorage, AK**

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(AKaerial)
library(kableExtra)
library(tinytex)
library(zoo)
library(tmap)
load(file="data/YKD_report.Rdata")
source("ObsMap2.R")
#IMPORTANT!!!  For first time users and/or to render the most up-to-date info, the user must first go to their R Console and source the YKD_report.r code. To do so, type 'source("YKD_report.R")' at the > prompt.  That will run the code to create all the species indices stats, study area map, table 2 & 3, and appendices.  Sourcing that will need to be done every time a change is made to that YKD_report.r code for rendering from YKD_report.qmd.  Rtools 4.3 may also need to be installed/reinstalled and can be found online or in FWS Apps-to-Go. Also, to insure the most recent version of AKAerial is being used, run the following code: devtools::install_github("USFWS/AKaerial", ref = "master", build_vignettes = TRUE). If you've already run devtools and are simply refreshing, use this code: devtools::install_github("USFWS/AKaerial", ref = "master", build_vignettes = FALSE).
```


# ABSTRACT {.unnumbered}

The Yukon Kuskokwim Delta Aerial Breeding Pair Survey provides data on the distribution, abundance, and trends of 22 waterbird species that nest in western Alaska, including primary population data for the threatened spectacled eider, emperor geese, cackling geese, Pacific greater white-fronted geese, and Pacific black brant from 1985 –`r max(YKDHistoric$combined$Year)`. No surveys were conducted in 2020 due to the COVID pandemic.  The most recent (`r max(YKDHistoric$combined$Year)`) indicated total population indices for spectacled eiders was `r formatC(SPEI$itotal[SPEI$Year==2024], big.mark=",", format="d")` (95% CI `r formatC(SPEI$itotal[SPEI$Year==2024]-1.96* SPEI$itotal.se[SPEI$Year==2024], big.mark=",", format="d")` - `r formatC(SPEI$itotal[SPEI$Year==2024]+1.96* SPEI$itotal.se[SPEI$Year==2024], big.mark=",", format="d")`), for emperor geese was `r formatC(EMGO$itotal[EMGO$Year==2024], big.mark=",", format="d")` (95% CI `r formatC(EMGO$itotal[EMGO$Year==2024]-1.96* EMGO$itotal.se[EMGO$Year==2024], big.mark=",", format="d")` - `r formatC(EMGO$itotal[EMGO$Year==2024]+1.96* EMGO$itotal.se[EMGO$Year==2024], big.mark=",", format="d")`), for cackling geese was `r formatC(CCGO$itotal[CCGO$Year==2024], big.mark=",", format="d")` (95% CI `r formatC(CCGO$itotal[CCGO$Year==2024]-1.96* CCGO$itotal.se[CCGO$Year==2024], big.mark=",", format="d")` - `r formatC(CCGO$itotal[CCGO$Year==2024]+1.96* CCGO$itotal.se[CCGO$Year==2024], big.mark=",", format="d")`), for Pacific greater white-fronted geese was `r formatC(GWFG$itotal[GWFG$Year==2024], big.mark=",", format="d")` (95% CI `r formatC(GWFG$itotal[GWFG$Year==2024]-1.96* GWFG$itotal.se[GWFG$Year==2024], big.mark=",", format="d")` - `r formatC(GWFG$itotal[GWFG$Year==2024]+1.96* GWFG$itotal.se[GWFG$Year==2024], big.mark=",", format="d")`), and for Pacific black brant was `r formatC(BRAN$itotal[BRAN$Year==2024], big.mark=",", format="d")` (95% CI `r formatC(BRAN$itotal[BRAN$Year==2024]-1.96* BRAN$itotal.se[BRAN$Year==2024], big.mark=",", format="d")` - `r formatC(BRAN$itotal[BRAN$Year==2024]+1.96* BRAN$itotal.se[BRAN$Year==2024], big.mark=",", format="d")`).  Long-term (`r min(EMGO$Year)`-`r max(EMGO$Year)`) mean annual population indices and trends (of indicated total  birds and their associated 95% CIs) for these species were as follows: spectacled eiders `r formatC(SPEIh, big.mark=",", format="d")` (`r formatC(SPEIh-1.96*SPEIhse, big.mark=",", format="d")` - `r formatC(SPEIh+1.96*SPEIhse, big.mark=",", format="d")`), growth rate `r sppdf$Trend[sppdf$Species == "SPEI"]`, emperor geese `r formatC(EMGOh, big.mark=",", format="d")` (`r formatC(EMGOh-1.96*EMGOhse, big.mark=",", format="d")` - `r formatC(EMGOh+1.96*EMGOhse, big.mark=",", format="d")`), growth rate `r sppdf$Trend[sppdf$Species == "EMGO"]`, cackling geese `r formatC(CCGOh, big.mark=",", format="d")` (`r formatC(CCGOh-1.96*CCGOhse, big.mark=",", format="d")` - `r formatC(CCGOh+1.96*CCGOhse, big.mark=",", format="d")`), growth rate `r sppdf$Trend[sppdf$Species == "CCGO"]`, Pacific greater white-fronted geese `r formatC(GWFGh, big.mark=",", format="d")` (`r formatC(GWFGh-1.96*GWFGhse, big.mark=",", format="d")` - `r formatC(GWFGh+1.96*GWFGhse, big.mark=",", format="d")`), growth rate `r sppdf$Trend[sppdf$Species == "GWFG"]`, and Pacific black brant `r formatC(BRANh, big.mark=",", format="d")` (`r formatC(BRANh-1.96*BRANhse, big.mark=",", format="d")` - `r formatC(BRANh+1.96*BRANhse, big.mark=",", format="d")`), growth rate `r sppdf$Trend[sppdf$Species == "BRAN"]`. Our population indices do not account for incomplete detection or other sources of bias at this time, although the model we used to estimate growth rate removes average observer effects. Posterior probabilities for a long-term (1985 –`r max(YKDHistoric$combined$Year)`) growth rate \< 1.00 were \> 0.975 for northern shoveler, American wigeon, northern pintail, green-winged teal, canvasback, long-tailed duck, and jaeger, indicating high confidence in long-term population declines for these species. By the same metrics, posterior probabilities for population increase were $\geq$ 0.90 for emperor geese, greater white-fronted geese, cackling and Taverner's geese, common eider, and red-breasted merganser, indicating moderate confidence in long-term population increases for these species. 
By the same metrics, posterior probabilities for population increase were $\geq$ 0.90 for emperor geese (0.97), greater white-fronted geese (1.00), cackling (1.00) and Taverner's geese (0.93), common eider (1.00), and red-breasted merganser (0.97), indicating moderate to high confidence in long-term population increases for these species. Growth rates for brant, tundra swans, mallards, scaup species, spectacled eiders, black scoters, and red-throated loon indicated relatively stable populations (95% credible interval bounds for long-term growth rate were $\pm$ 4% of a long-term growth rate of 1.00) and/or confidence in directional trends that was more limited (\< 0.90 posterior probability for decrease or \< 0.80 for increase).  In addition to long-term trends, we examined more recent (10-year) patterns of abundance for all species. Herein, upper 95% credible intervals for the most recent 10-year (2014 – 2024) growth rate for greater white-fronted geese, brant, cackling geese, American wigeon, northern pintail, spectacled eider, sandhill crane, jaegar, and Pacific loon were all \< 1.00 and the posterior probability that the trend was decreasing was \> 0.975, indicating high confidence in more recent population decreases for those species.  Other species showed less evidence (lower posterior probability) for a population decrease, but still demonstrated relatively high confidence in a decreasing trend, included emperor geese (0.97), tundra swans (0.96), and green-winged teal (0.97). 
There was only moderate confidence in population increase across the most recent 10-years (based on posterior probabilities $\geq$ 0.80), for red-breasted mergansers (0.90) and common eiders (0.85). Most recent 10-year rowth rates for northern shoveler, mallards, black scoters, and red-throated loons indicated relatively stable populations (95% credible interval bounds for long-term growth rate were $\pm$ 4% of a long-term growth rate of 1.00) and/or confidence in directional trends that was more limited (\< 0.90 posterior probability for decrease or \< 0.80 for increase). Evidence of more recent (10-year) negative population trajectories for species of concern, such as emperor geese and spectacled eiders, deserves continued attention.  

## *Key Words* {.unnumbered}

Aerial breeding pair survey, Alaska, Emperor goose *Chen canagica*, population index, growth rate, species occurrence, spectacled eider *Somateria fischeri*, waterfowl, waterbird, Yukon-Kuskokwim Delta

# INTRODUCTION {.unnumbered}

The coastal zone of the Yukon-Kuskokwim (YKD) is a vast area of sub-arctic lowland tundra in far western Alaska, important for many species of breeding birds [@Gill_Handel1990; @SaalfeldYKD2017; @Fischer2018; @Jorgenson2018]. Importantly, it is the primary breeding area in Alaska for emperor geese [*Chen canagica* holding ~80% of the world's population; @Eisenhauer1977], cackling geese [*Branta hutchinsii minima*; supporting all of the subspecies, @King1981], Pacific greater white-fronted geese [*Anser albifrons frontalis*; supporting 95% of the population, @Timm1979], and is one of only three breeding areas worldwide for the threatened spectacled eider [*Somateria fischeri*: @USFWS1996].  Additionally, the YKD hosts many other waterbird species of conservation concern, including red-throated loons (*Gavia stellata*), Pacific black brant (*Branta bernicla nigricans*), common eiders (*S. mollissima*), black scoters (*Melanitta nigra*), and long-tailed ducks (*Clangula hyemalis*). Significant waterfowl population declines on the YKD from the 1960's to 1980's [@King1981; @Stehn1993YKD; @Ely1994] resulted in restrictions to harvest for geese [@Pamplin1986], and listing of the spectacled eider as threatened under the Endangered Species Act [@USFWS1993; @USFWS1996]. 

These declines led to development of annual aerial and ground-based surveys of waterbirds on the YKD [@Stehn1993YKD; @Fischer2018], and formal monitoring efforts to measure population indices and trends were instituted in 1985 [@Fischer2018].  Herein, we present results from data collected on the YKD Aerial Breeding Population Survey (hereafter "YKD Survey") from `r min(YKGHistoric$combined$Year)`-`r max(YKGHistoric$combined$Year)`, providing distribution, abundance, and trend information for 22 species of birds, including geese, swans, cranes, ducks, and jaegers.  These results provide population status and distribution information to the Pacific Flyway Council, Alaska Migratory Bird Co-Management Council, Spectacled Eider Recovery Team, U.S. Fish and Wildlife Service, U.S. Geological Survey, Association of Village Council Presidents, and various non-governmental organizations.  Our report focuses on the previously unreported 2017-2024 survey data and presents updated population information for all 22 bird species monitored from `r min(YKGHistoric$combined$Year)`-`r max(YKGHistoric$combined$Year)`.

# METHODS {.unnumbered}

## Survey Design {.unnumbered}

The YKD Survey study area extends along the Yukon-Kuskokwim Delta coast from Kuskokwim Bay in the south to Norton Sound in the north, and from the coastal fringe in the west to approximately 50 km (31 mi) inland (@fig-Fig1). The coastal zone was originally divided into 16 strata based on homogeneous physiographic regions determined from unclassified LANDSAT imagery (Butler 1988). The survey design was reduced to four primary strata based on goose densities in 1998 (@Platte2009). In 2004, the current 5-strata design was developed and an additional stratum was created to accommodate historic data in a high-density area where transect spacing varied over time (@fig-Fig1) . In developing the sampling scheme, a randomly selected starting point was used to construct a series of systematically placed transects in an east-west orientation. Transects were spaced at 1.6 km (1.0 mi) intervals in the highest density stratum, and at 3.2 km (2.0 mi), 6.4 km (4.0 mi), and 12.9 km (8.0 mi) intervals in high, medium, and low density strata, respectively (@fig-Fig1). Since 1998, this aerial survey has used a rotating panel design to balance the spatial coverage of transects for estimating population means with temporal replication for monitoring trends over time. Near complete coverage (>95%) of the highest density stratum was obtained over a four-year rotation by shifting transects 0.4 km each year.  

## Survey Methods {.unnumbered}

The YKD Survey methodology is based on standard operating procedures for the North American Waterfowl Breeding Population and Habitat Survey (USFWS and CWS 1987), with both the left-front seat biologist-pilot and right-front seat observers recording all swans, geese, and cranes, and a single rear-seat observer recording all other species (i.e., loons, ducks, grebes) within a strip transect encompassing 200 m either side of the flight path.  To estimate the outer transect boundary, crews determine the required viewing angle trigonometrically and mark a reference point on the wing strut and side window for each observer, using a clinometer and marking tape/pens. 
Surveys are flown at an approximate ground speed of 161 km/hr (~87 kts; though actual ground speeds vary due to winds), and an altitude of 38 m (125 ft) above ground level (AGL), as referenced by a radar altimeter installed in the aircraft. Georeferenced observations made from both sides of the aircraft are voice-recorded into panel-mounted devices for later transcription using the custom software packages RECORD and TRANSCRIBE [2018-2022: @Hodges2015Record] and H2 and SCRIBE [2023-present: @H2].  The survey platform in all years was an amphibious or float-equipped Cessna 206 aircraft (Cessna Aircraft Company, Wichita, KS), with the exception of 2012, when an amphibious Kodiak 100 was used (@tbl-crew). 


## Survey Timing and Weather Conditions {.unnumbered}

This survey was timed to coincide with the egg laying / early incubation period of nesting geese and eiders on the YKD, and the peak presence of male ducks and pairs of other waterfowl species.  This is a period when nesting habitat is available (i.e., ice-free water is visible in most shallow vegetated wetlands, and tundra vegetation is mostly snow-free), typically within the last two weeks of May through early June. Previously-collected nest plot data from the YKD coastal zone [@Fischer2018] were used to select a general target date, which was refined using Meteorological Terminal Air Report (METAR) temperature data (available online at <http://www.wunderground.com/>). Exact start dates were predicted based on the long-term (1993–2015) relationship between spring warming temperatures and the average annual clutch initiation date for cackling Canada geese. Air temperature data from weather stations in Bethel, Cape Romanzof, Emmonak, Hooper Bay, Mekoryuk, and St Marys were used to calculate the number of accumulated thaw-degree-days (TDD) where daily mean air temperatures exceeded 32°F, and the average TDD was >25 days at all six sites [@PlatteStehn2015, @Swaim2017]. Based on the relationship between these temperature deviates and the predicted nest initiation dates derived from the YKD nest plot study (Fischer et al. 2016), surveys were scheduled to begin nine days after the predicted nest initiation date for cackling Canada geese. Within the last two weeks of May each year (@tbl-crew), we refined our survey start dates based on close monitoring of weather and temperature data, examining snow and ice-cover changes via satellite ([NASA Worldview](https://worldview.earthdata.nasa.gov/)) and web-camera imagery (<https://weathercams.faa.gov//>), and by receiving updates on current landscape conditions from biologists and other local residents on the YKD. 
From 2017–2024, weather conditions varied considerably within and among survey years. Reduced visibility and ceilings due to coastal fog, high winds, and turbulence were the most consistent weather impediments to surveying in all years of the survey.

## Population Indices {.unnumbered}

We calculated population indices in a fashion to be consistent with their use in other USFWS surveys. Duck indices followed the guidance of @USFWSandCWS1987, and goose, swan, and crane indices followed @Eldridge2003. Pairs were defined as the total number of male-female pairs (or two monomorphic birds \[e.g., geese\] in close association), not the total number of birds in pairs [@USFWSandCWS1987]. Flocked drakes were defined as the total number of males observed with one or more other males and no female present [@USFWSandCWS1987], and were only recorded for ducks. Flocked drakes in groups of less than 5 birds were doubled (except for scaup; see below), and flocked drakes in groups of 5 or more were treated as a flock and not doubled [@USFWSandCWS1987]. Flocks were defined as a closely associated single-sex or mixed-sex grouping of 5 or more birds (or 2 or more birds of different species) that could not be separated into singles and pairs. For scaup, single drakes and flocked drakes in groups of less than 5 birds were not doubled as they were for other ducks. This is because sex ratios in scaup lean heavily towards males, such that not all males can be assumed to have a female mate [@USFWSandCWS1987]. Formulas for the calculation of the four population indices presented in this report are shown below.


{{< pagebreak >}}

![Study area and survey design for the Yukon-Kuskokwim Coastal Zone Breeding Population Survey (YKD Survey) in western Alaska, illustrating the 2024 transects, strata, and village locations. The exact locations of survey transects vary slightly each year based on a four-year rotating panel within four physiographic and bird-density-based strata (very high, high, medium, low).](data/fig1.png){#fig-Fig1 fig-alt="Study area and survey design map for the Yukon-Kuskokwim Coastal Zone Aerial Breeding Population Survey (YKD Survey) in northern Alaska, illustrating the 2024 transects, strata, and village locations. The exact locations of survey transects vary slightly each year based on a four-year rotating panel within four physiographic and bird-density-based strata (high, medium, low, and very high)."}

{{< pagebreak >}}

```{r table1, include=TRUE, echo=FALSE}
#| label: tbl-crew
cap <- paste0 ("Yukon-Kuskokwim Delta Breeding Population Survey (YKD Survey) dates and crews (1985-present). The current YKD Survey combines counts of swans, geese, and cranes from the two front-seat observers (1985-present), with data from a single rear-seat observer recording all other species (i.e., loons, ducks, grebes; 1988-present). No rear-seat observer was present in 2011, and pilots (Bart Stone and Mike Callahan, respectively) in 2012-2013 did not collect data. No surveys were flown in 2020 due to the COVID pandemic. The survey platform in all years was an amphibious or float-equipped Cessna 206 aircraft, with the exception fo 2012, when an amphibious Kodiak 100 was used.")

kable(table1, digits = 0, booktabs = TRUE, linesep = "", longtable=TRUE, caption=cap, align = "cccccc", col.names = c("Year", "Data Collection", "Left-front", "observer", "observer", "Aircraft")) %>%
  kable_styling(font_size=10, position="center") %>%
  add_header_above(bold = TRUE, c(" "=1, "Dates of" = 1, "Observer & Pilot" = 1, "Right-front" = 1, "Right-rear" = 1, " "=1)) %>% 
  row_spec(row = 0, bold = TRUE) 

```


For dimorphic species (e.g., ducks \[except scaup\]) and some monomorphic species (cranes and dark geese \[e.g., greater white-fronted geese, cackling/Canada geese, and brant\]):

```{=tex}
\begin{center}

\textit{Indicated Breeding Birds = 2(singles + pairs + flocked drakes in groups of < 5 birds)} \newline

\textit{Indicated Total Birds = 2(singles + pairs + flocked drakes in groups of < 5 birds) + birds in flocks} \newline

\end{center}

For the remainder of the monomorphic species (e.g., swans, snow geese, grebes, loons, terns, gulls, jaegers, owls, eagles, ravens) and scaup species:\newline

\begin{center}

\textit{Breeding Birds = singles + 2(pairs) + flocked drakes in groups of < 5 birds} \newline

\textit{Total Birds = singles + 2(pairs) + flocked drakes in groups of < 5 birds + birds in flocks} \newline

\end{center}
```

## The Ratio Estimator {.unnumbered}
Population totals for each index were calculated using a ratio estimator [@Cochran1977] with transects as the sample units. The classical ratio estimator [@Cochran1977] is a straightforward approach that can be used to estimate the population total of observed birds in a survey area given a systematic or random sampling design with large sample size (i.e., number of transects). The ratio estimator is especially good when the response (i.e., number of birds on a transect) is linearly related to the length or area of the transect. We employed a stratified strip-transect design for the majority of our surveys of breeding waterfowl. Study areas are separated into spatial strata based on physiographic features and historical bird densities. Strata were sampled by low-level flights along a series of strip transects, with one or two observers searching 200 meters out from the transect center. Observations and sampled areas are then summed across transects within each stratum to produce strata-specific density estimates. These densities are then multiplied by associated strata areas to produce stratum-specific population indices that are summed to produce the overall population indices, $$E[\hat Y] = \displaystyle\sum_{i}^{S} \frac{\bar y_i}{\bar a_i} A_i = \displaystyle\sum_{i}^{S} \hat D_i A_i$$ where $y_i$ are observations and $a_i$ are sampled areas in strata $i$, and $$A_i = \displaystyle\sum_{j}^{M_i} a_{ij}$$ with estimated variance from Williams et al. (2002), where standard error (SE) is calculated based on inter-transect variance,  $$Var(\hat Y) = \displaystyle\sum_{i}^{S} M_i^2 \frac{(1-m_i / M_i)}{m_i} ( s^2_{iy} + \hat {D^2}_i s^2_{ia} - 2 \hat {D_i} s_{iay} ),$$ where $m_i$ and $M_i$ are the number of sampled and total transects in strata $i$, respectively, and there are $S$ total strata;  $$S^2_{ix} = \displaystyle\sum_{j}^{m_i} (x_{ij} - \bar x )/(m_i -1), $$ and  $$S^2_{ixy} = \displaystyle\sum_{j}^{m_i} (x_{ij} - \bar x )(y_{ij} - \bar y )/(m_i -1).$$ The equations above are observer-specific (i.e., calculated for each observer independently). However, the estimates presented in this report are the arithmetic mean of observer-specific estimates (n=2, in most years). Design transects (@fig-Fig1) were used to calculate survey effort, but flight lines sometimes deviated slightly from the these due to weather avoidance or other factors. Data manuplulation was completed using the R [@R] package `AKaerial` [@r-akaerial].

Population indices presented here do not account for incomplete detection, though efforts to quantify and incorporate detection rates into annual population estimates are being developed [@Lewis2019; @Osnas2024]. Summary statistics were calculated for the indicated total bird (or total bird) index for each bird species, including the long-term (`r min(SPEI$Year)`- `r max(SPEI$Year)`) and most recent 3-year averages (2022-2024) and their standard errors (@tbl-results2). Indicated breeding bird (or breeding bird) indices and indicated total bird (or total bird) indices for each bird species are presented in the Appendices (@tbl-EMGO - @tbl-PALO and @fig-EMGO - @fig-PALO). Throughout, we report point estimates $\pm$ (1.96 x SE) as a 95% confidence interval, except in the simulations for trends (see next section).

## Index Trends {.unnumbered}

Trajectories in population indices were estimated using generalized additive models [GAM, @wood2017] in the R package `mgcv` [@r-mgcv] fit to observer-specific index estimates. The GAM model used a scaled-t likelihood, a continuous smooth term for year, and a random effect for observer in the linear predictor. The observer effect allowed us to remove observer-specific effects from any estimated average trajectory, and the scaled-t likelihood allowed for extra residual variance from the trend line over a normal likelihood that might be due to additional year-specific variance from the trend line. To account for uncertainty in the index estimate, we used a parametric bootstrap of the estimated index as the modeled response in the GAM. Thus, for each year, a response was sampled from a normal distribution (with the mean and standard deviation of the point estimate), truncated to a small value (1e-10) for samples $\leq$ 0, and log-transformed. Then a GAM was fit, and a sample of the parameters was obtained from a multivariate normal distribution using the estimated parameter vector and covariance matrix. Annual predictions over the full time series were made from this parameter sample after removing the effect of observer, and the results were saved. This procedure was repeated 300 times. As such, this procedure is an empirical Bayesian method to approximate the posterior distribution of the trajectory [@miller2025bayesian]. For species which were not observed in all years (e.g., red-breasted mergansers), we excluded years when no birds were observed. We did this because observed zeros in these design-based estimates are more likely due to non-detection or random sampling, rather than complete absence of the species in the survey area, and model-fit to data that included zeros did not produce stable estimates. However, we acknowledge that this may have resulted in some over-smoothing and bias of estimated trends for species with zero-count years. For improved methods, that better account for zero observations in rare species see [@Amundson2019, @Osnas2024].


We summarized the simulated yearly GAM predictions (i.e., the posterior distribution of trajectories) by calculating the mean and standard deviation for each year. An example of simulated results and observer-specific estimates for one species (greater white-fronted goose) is shown in @fig-gwfggam. For plots and tables presented in the Results and Appendices, we do not report the observer-specific index estimates or individual samples of the smooths, in an effort to reduce clutter. Instead, we show only the average index between observers in any year, the mean posterior trajectory, and $\pm$ 2 standard errors from this mean.

![An example of the modeled index estimates and trend results for greater white-fronted geese, accounting for observer effects. Colored points and bars are observer-specific index estimates, which were sampled 300 times and used as the response in a generalized additive model. Individal modeled predictions for each sample are shown by the thin gray lines. These make up the posterior distribution of modeled trajectories and are summarized by the orange band (approximate 95% credible interval) and the thick gray line (posterior mean). End points of the modeled predictions (thin gray lines) are used to calculate the posterior distribution of the long-term growth rate.](data/plot_trends/figures/GWFG_index_2.png){#fig-gwfggam width="50%" height="50%" fig-alt="A figure showing observer-specific estimate for greater white-fronted geese and 300 posterior samples of the modeled GAM trajectories, the posterior mean trajectory, and the approximate 95% confidence interval on the trajectory."}

We estimated the long-term average growth rate from year $a$ to $b$ for each simulated replicate as

$$T_{b-a} = (\hat N_{b}/\hat N_{a})^{1/(b-a)},$$
where $\hat N_t$ is the predicted index value at year $t$ from the bootstrap replicate GAM model, $t = a$ is the first year of a non-zero estimate, and $t = b$ is the last year. Thus, the growth rate was calculated as the geometric mean annual change in the model-predicted index. We summarized growth rate by the mean and 95% credible interval across the sampled replicates. We also calculated the posterior probability of a negative growth rate by summarizing the proportion of posterior samples that were \< 1 (Probability of Trend < 1, @tbl-results3). We only calculated growth rate for total or indicated total bird indices shown in @tbl-results3. Code is available in the file `plot_trends_design.R` at [@osnasgithub].

# RESULTS {.unnumbered #sec-results}

## Survey Timing and Weather Conditions {.unnumbered}

The 2024 survey represents the first year of the seventh 4-year rotation of this survey, during which 104 transects were flown, covering 2,047 linear km. Dates of data collection for 2017-2019 and 2022-24 are shown respectively in (@tbl-crew). The surveys required 32, 34, 35, 41, 38, 37, and 34 flight-hours to complete, respectively, 2019 and 2021-2024, not including ferry time to and from the study area.  Survey timing was the earliest on record for the survey in 2019 (23 May - 5 June; @tbl-crew). The first day of data collection on the YKD in 2019 was 8 days earlier than the (2005-2024) average start date of 31 May. In 2019, there was little to no snow or ice on the landscape over the late winter and spring and warm temperatures resulted in an early break-up.  

{{< pagebreak >}}

In contrast, 2024 marked a later year, with sustained cold temperatures and significant snow and ice covering much of the coastal YKD landscape into late May, as well as a mid-survey snowfall in early June. Despite the later phenology of 2024, 2018-2023 (all with May start dates) were relatively early years, compared to historical surveys, which typically started in June (@tbl-crew). 



Though not a focus of our report, we conducted exploratory analyses of possible effects of nesting phenology and survey timing on the estimates from the YKD aerial survey. Herein, we examined the difference between the within-year median day of the survey and the median day of nest initiation, as measured by egg floating from the YKD Nest Plot Survey [@Fischer2018] or distance sampling surveys (Osnas pers. comm). Our preliminary results revealed minimal effects of timing. However, we did find that the YKD Survey has been getting earlier over time (likley in response to earlier nest initiation), but also *relative* to initiation. In other words, the surveys appear to be appropriately adjusted to phenologically driven changes in timing of nesting, but also tend to be occurring earlier in the nesting cycle despite early or late years (i.e., slightly earlier in initiation/early incubation), relative to historical surveys.  The result of this shift has likely resulted in the survey being more "optimally timed" for the availability of breeding male duck species (i.e., before they depart the breeding grounds) and pairs (i.e., before nest failure). 

## Population Indices {.unnumbered}

We summarize the most recent 3-year average (2022-2024) and long-term average (`r min(YKGHistoric$combined$Year)`-`r max(YKGHistoric$combined$Year)`) population indices, as well as long-term (`r min(YKGHistoric$combined$Year)`-`r max(YKGHistoric$combined$Year)`) posterior mean trajectories (trends) and associated 95% credible intervals (CI, @tbl-results2), as well as posterior probabilities that trends are decreasing (Probability of trend < 1) for 22 bird species observed on the YKD Survey (@tbl-results3). Herein, we follow the taxonomic ordering of @Chesser2024. For each species, we also present individual annual population indices (`r min(YKGHistoric$combined$Year)`-`r max(YKGHistoric$combined$Year)`) and trajectories in graphical and tabular formats, and maps of individual species-observations for 2024 in the the Appendices (@tbl-EMGO - @tbl-PALO and @fig-EMGO - @fig-PALO). Though our maps provide accurate observation locations for each species, we caution that our location data should not be interpreted as densities, due to differences in sampling effort across strata boundaries. Historical YKD densities have been presented by @Platte2009 and @SaalfeldYKD2017, but more recent density-modeling efforts for YKD species have not yet been completed [though see @Osnas2024 for an example with Steller's eiders on the Arctic Coastal Plain].

```{r table2, include=TRUE, echo=FALSE}
#| label: tbl-results2

#TABLE 2 - Creates a summary table with a row of estimates for each species

cap <- paste0 ("The most recent 3-year (2022-2024) and long-term (",min(YKDHistoric$combined$Year),"-",max(YKDHistoric$combined$Year),") average population indices and associated standard errors (SE), as well as most recent 10-year and long-term annual growth rates (long-term trend), associated 95% credible intervals (CI), and posterior probability that each trend is decreasing (Probability of trend < 1]) for bird species on the Yukon-Kuskokwim Delta Aerial Breeding Population Survey in Alaska.  Index formulas used are species-specific, with indicated total birds (ITB) used for dimorphic species, dark geese, and sandhill cranes, and total birds (TB) used for the remaining monomorphic species, other than dark geese, sandhill cranes, and scaup. See the Methods section for detailed index definitions and trend calculations. Estimates are not corrected for incomplete detection, but trends do account for observer effects.")

kable(table2, digits = 0, booktabs = TRUE, longtable=TRUE, 
      format.args = list(big.mark = ","), caption=cap, align = "llrrrr", 
      col.names = c("Species", "Index", "Average", "SE", "Average", "SE"), 
      linesep = "") %>%
  kable_styling(font_size=10, position="center") %>%
  add_header_above(bold = TRUE, c(" "=1, " "=1, "3-Year" = 1, " "=1, 
                                  "Long-term" = 1, " "=1)) %>%
  row_spec(row = 0, bold = TRUE)

```

```{r table3, include=TRUE}
#| label: tbl-results3
#TABLE 3 - Creates a summary table with 10-yr and long term trend estimates for each species

cap <- paste0 ("The most recent 10-year (2014-2024) and long-term (",min(YKDHistoric$combined$Year),"-",max(YKDHistoric$combined$Year),") average population indices and associated standard errors (SE), as well as most recent 10-year and long-term annual growth rates (long-term trend), associated 95% credible intervals (CI), and posterior probability that each trend is decreasing (Probability of trend < 1) for bird species on the Yukon-Kuskokwim Delta Aerial Breeding Population Survey in Alaska.  Index formulas used are species-specific, with indicated total birds (ITB) used for dimorphic species, dark geese, and sandhill cranes, and total birds (TB) used for the remaining monomorphic species (other than dark geese, sandhill cranes) and scaup. See the Methods section for detailed index definitions and trend calculations. Estimates are not corrected for incomplete detection, but trends do account for observer effects.")

kable(table3, digits = 0, booktabs = TRUE, longtable=TRUE, 
      format.args = list(big.mark = ","), caption=cap, align = "llcccc", 
      col.names = c("Species", "Index", "Trend (CI)", "Trend < 1", "Trend (CI)",
                    "Trend < 1"), linesep = "") %>%
  kable_styling(font_size=10, position="center") %>%
  add_header_above(bold = TRUE, c(" "=1, " "=1, "10-year"=1, "Probability"=1,
                                  "Long-term"=1, "Probability"=1)) %>%
  row_spec(row = 0, bold = TRUE)

```


# DISCUSSION {.unnumbered}

This report describes trends in relative abundance for all common waterfowl and sandhill cranes on the YKD of Alaska, collected by USFWS from `r min(YKGHistoric$combined$Year)`-`r max(YKGHistoric$combined$Year)`; including 2017–2024 survey data not previously reported. Upper 95% credible intervals for long-term (`r min(YKGHistoric$combined$Year)`–`r max(YKGHistoric$combined$Year)`) growth rate for northern shoveler, American wigeon, northern pintail, green-winged teal, canvasback, long-tailed duck, jaeger were all \< 1.00 and the posterior probability that the trend was decreasing was \> 0.975, indicating high confidence in population decreases for those species.  Other species showed less evidence (lower posterior probability) for a population decrease (@tbl-results3). With less, but still relatively high confidence, sandhill crane (0.89), and Pacific loon (0.96) also showed evidence for a decreasing trend (@tbl-results3). 

By the same metrics, posterior probabilities for population increase were $\geq$ 0.90 for emperor geese (0.97), greater white-fronted geese (1.00), cackling (1.00) and Taverner's geese (0.93), common eider (1.00), and red-breasted merganser (0.97), indicating moderate to high confidence in long-term population increases for these species. Growth rates for brant, tundra swans, mallards, scaup species, spectacled eiders, black scoters, and red-throated loon indicated relatively stable populations (95% credible interval bounds for long-term growth rate were $\pm$ 4% of a long-term growth rate of 1.00) and/or confidence in directional trends that was more limited (\< 0.90 posterior probability for decrease or \< 0.80 for increase; @tbl-results3).

In addition to long-term trends, we examined more recent (10-year) patterns of abundance for all species. Herein, upper 95% credible intervals for the most recent 10-year (2014 – 2024) growth rate for greater white-fronted geese, brant, cackling geese, American wigeon, northern pintail, spectacled eider, sandhill crane, jaegar, and Pacific loon were all \< 1.00 and the posterior probability that the trend was decreasing was \> 0.975, indicating high confidence in population decreases for those species.  Other species showed less evidence (lower posterior probability) for a population decrease (@tbl-results3), but still demonstrated relatively high confidence in a decreasing trend, included emperor geese (0.97), tundra swans (0.96), and green-winged teal (0.97; @tbl-results3). Evidence of more recent negative population trajectories for species of concern, such as emperor geese and spectacled eiders, deserves careful monitoring.

There was only moderate confidence in population increase across the most recent 10-years (based on posterior probabilities $\geq$ 0.80), for red-breasted mergansers (0.90) and common eiders (0.85). Growth rates for northern shoveler, mallards, black scoters, and red-throated loons indicated relatively stable populations (95% credible interval bounds for long-term growth rate were $\pm$ 4% of a long-term growth rate of 1.00) and/or confidence in directional trends that was more limited (\< 0.90 posterior probability for decrease or \< 0.80 for increase; @tbl-results3).

Observer bias is a factor in all aerial surveys. However, over the long-time span of the YKD Survey (38 years of data collection between 1985-2024), there have been only 6 left-seat pilot obervers, 8 right-front seat observers (3 of whom served as pilot-observers in other years), and 5 right-rear seat observers. To account for observer bias, we used a model to statistically remove average observer effects from the estimated average trajectory. This represents an improvement over past long-term trend estimates, and lacking any trend in detection or other observation biases, the trends reported here are the best currently available for this area without directly estimating these observer biases. Nevertheless, detection and availability biases are still largely unaccounted for, affecting our population indices. Thus, these estimates are not measures of absolute abundance. Though we do not provide detection-adjusted population estimates for the YKD Survey, studies examining observer detection of spectacled eiders on the YKD have been conducted [@Lewis2019] and we hope to incorporate detection and observer-adjusted estimates for all species into future reports, using methods similar to @Osnas2024.

The trajectories estimated here represent continuous smooth functions after removing average observer effects. We did not fit other models that allowed for more complex or discontinuous trends, such as year-specific random effects or spatially explicit predictions [e.g. @smith2021north; @Osnas2024]. Because many species do show large year-specific deviations from an average trend, a model with such effects is likely a better description for these species. Such models are not well-estimated when fit to the design-based estimates used here. The model we have used, however, can be viewed as a flexible way to estimate the average trajectory through time, smoothing over year-specific variations. Compared to a simple log-linear regression on point estimates, the current approach provides for more flexible functional forms of the trajectory, allowing for the identification of cycles or periods of change. One should keep these points in mind when interpreting results, as for many species, the year-specific variations are much larger than the average change over time (e.g., mallards) or show potentially cyclical patterns over a shorter time (e.g., sandhill cranes).

Trajectories for rare species (e.g., red-breasted mergansers; those with observed zero estimates) should be interpreted with care. We attempted to fit our model by including zero estimates (after adding a small number and log transforming) but the GAM often failed to find stable estimates, so we used only positive estimates. This will have two important consequences. First, trajectories may be biased if there is a trend in the frequency of zero observations in the time series. Second, the GAM will tend to produce smoother trajectories than if the zeros were included. For example, the red-breasted merganser trajectory is essential a flat line (@fig-RBME), though more subtle, cyclical patterns could be possible. Interestingly, @Amundson2019 also excluded zero-counts in a spatially-explicit GAM for the Arctic Coastal Plain Aerial Survey, and found a flat trajectory for Steller's eider over a much longer time period; suggesting that the smooth GAM is due to a reduction in data when most observations are excluded. We expect similar patterns for other rare species. However, for more common species that do not include zero estimates, we believe the current model should approximate temporal trajectories in relative abundance. In the future, we also hope to fit a spatial model to the full set of species.

While several targeted monitoring programs for specific species or limited areas of the YKD Coastal Zone have been conducted; to our knowledge, the YKD Survey represents the most comprehensive aerial survey of sub-arctic waterbirds in existence. The analysis and implementation of a multi-species, aerial monitoring program such as the YKD Survey is challenging due to the inherent variability of the natural ecosystem, the varied natural histories of 22+ species, and the logistical difficulties of conducting aerial surveys in the Alaska. Moreover, it is difficult to achieve adequate within-year sampling of such a large spatial area within a short annual phenological window. Surveys of this type are further complicated by several waterfowl species that exist in extremely low densities (such as canvasbacks and red-breasted mergansers), making precise estimates and trends difficult to achieve [though see @Osnas2024 for an example with Steller's eiders]. Collection of such a long-term data set has been made possible by integrating improvements in sample design along the way. The sample design used in this survey was originally developed in 1985 to target breeding geese [@Fischer2018], and later augmented with a third (rear-seat) observer to simultaneously monitor breeding waterbirds beyond geese, swans, and cranes [@Fischer2018]. The current survey provides good temporal, spatial, and inferential compromise. However, the changing distributions and abundances of many waterfowl species, as well as the rapid and wide-spread landscape changes in the sub-arctic, warrants reevaluation of the current stratification and overall survey design. Nonetheless, this survey represents one of few broad-scale, long-term (38-year), systematic monitoring efforts for waterbirds in the sub-arctic of North America.

## *Supplemental Material* {.unnumbered}

Species-specific population indices, trends, and spatial distributions of observations can be found in the tables and figures of the Appendices. Original data for the YKD Survey can be found at Science Base - Alaska Yukon-Kuskokwim Delta Aerial Breeding Pair Survey <https://www.sciencebase.gov/catalog/item/662858ead34ea70bd5f056a5> and point-estimates (and documentation of their calculation) were sourced from the R package AKaerial at <https://github.com/USFWS/AKaerial>. R Quarto code used to produce this report, the report itself, and associated tabular data, can be found at <https://www.sciencebase.gov/catalog/item/6628594ad34ea70bd5f056fa>.

## *Suggested Citation* {.unnumbered}

Wilson, H.M., Swaim, M. A., Safine, D.E., Frost, C.J., Zeller, T.K., and E.E. Osnas. 2025. Population indices, trends, and distribution of breeding waterbirds on the Yukon-Kuskokwim Delta, Alaska, 1985-2024. Anchorage, Alaska: U.S. Fish and Wildlife Service, Migratory Bird Management.<https://doi.org/10.7944/f502-9171>.


# ACKNOWLEDGEMENTS {.unnumbered}

This project would not have been possible without the support of many individuals and organizations. We thank the staff at Yukon Delta National Wildlife Refuge (YDNWR) for their continued support and dedication to this aerial survey, with special thanks to Charles Rodgers (Aircraft Mechanic, YDNWR), who over the years has cheerfully responded to various maintenance requests. Many thanks also to the Alaska Department of Fish and Game and the USGS Alaska Science Center for periodic fuel support. We are grateful to Julian Fischer for his thoughtful report reviews, project support, and program leadership. Finally, we acknowledge the tremendous contribution of the observers and pilots who collected aerial waterbird data on the YKD beginning in 1985 through the present (@tbl-crew), and all those who supported their efforts.

<center>

![Yukon-Kuskokwim Delta Survey crews with amphibious-equipped Cessna 206 survey aircraft. Left to right, crew 1 (2017-2022): Heather Wilson (left-front seat observer/pilot), Tamara Zeller (right-rear seat observer), and Michael Swaim (right-front seat observer) with N375F at Kanagyak Field Station, and crew 2 (2023-2024): Michael Swaim, Heather Wilson, David Safine (right-front seat observer) with aircraft N721NR in Bethel, Alaska.))](data/Combined_YKD_Crew_pics.jpg){#fig-crew width="100%" height="75%" fig-alt="Photos of the Yukon-Kuskokwim Delta Survey crews with amphibious-equipped Cessna 206 survey aircraft. Left photo is of crew 1 (2017-2022): Heather Wilson (left-front seat observer/pilot), Tamara Zeller (right-rear seat observer), and Michael Swaim (right-front seat observer) with aircraft N375F at Kanagyak Field Station, and right photo is of crew 2 (2023-2024): Michael Swaim, Heather Wilson, David Safine (right-front seat observer) with aircraft N721NR in Bethel, Alaska."}  

</center>

# LITERATURE CITED {.unnumbered}

::: {#refs}
:::

{{< pagebreak >}}

{{< include _appendices.qmd >}}
